Drought vulnerability of tree growth

### 2024-07-10 scripts/explore_analysis.Rmd
updated resistance values for top 4 species
uses clean dataset

```{r plot increments}

@@

resistance_top4 <- top4 %>%
  filter(Cno %in% c(5, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(
    yr=ifelse(Cno==5, 2010, 2015),
        resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-avg_inc,
         resistance.prop = (inc_annual-avg_inc)/avg_inc)%>%
  filter(!is.infinite(resistance.div))

@@

quantiles_2010_top4<-quantile(resistance_2010_top4$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)
quantiles_2015_top4<-quantile(resistance_2015_top4$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

#split the data into categories based on quantiles

resistance_2010_top4$quantile<-ifelse(resistance_2010_top4$resistance.div<quantiles_2010_top4[1], 1, ifelse(resistance_2010_top4$resistance.div<quantiles_2010_top4[2], 2, ifelse(resistance_2010_top4$resistance.div<quantiles_2010_top4[3], 3, 4)))
resistance_2015_top4$quantile<-ifelse(resistance_2015_top4$resistance.div<quantiles_2015_top4[1], 1, ifelse(resistance_2015_top4$resistance.div<quantiles_2015_top4[2], 2, ifelse(resistance_2015_top4$resistance.div<quantiles_2015_top4[3], 3, 4)))

@@

resistance_div<-ggplot(data = resistance_all, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  facet_grid(.~yr)+
  xlim(-5, 5)+
  xlab("inc/long-term increment")+
  geom_vline(xintercept = c(0,1), lty=2)+
  theme_bw()

@@

resistance_div_sp<-ggplot(data = resistance_all, 
aes(x=resistance.div, 
col=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_density(alpha=0.5) +
  facet_grid(factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))~yr)+
  scale_color_viridis_d()+ scale_fill_viridis_d()+
  guides(col="none", fill="none")+
  xlim(-5, 5)+
  xlab("inc/long-term increment")+
  geom_vline(xintercept = c(0,1), lty=2)+
  theme_bw()

@@

resistance_map<-ggplot(resistance_all, aes(x=X, y=Y, size=as.numeric(DBH.04), col=quantile))+
  geom_point(alpha=0.5)+
   scale_size(range = c(.1, 10), name="DBH (cm)")+
  guides(color=guide_legend(title="resistance"))+
  scale_colour_gradient(low="red", high="blue") +
  facet_grid(sp~yr)+
  theme_minimal()+
  coord_fixed()

>>>>

```{r plot increments old code, echo=F, eval=F}

@@

resistance_top4 <- top4_noouts %>%
  filter(Cno %in% c(5, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(
    yr=ifelse(Cno==5, 2010, 2015),
        resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-avg_inc,
         resistance.prop = (inc_annual-avg_inc)/avg_inc)%>%
         ungroup()%>%
  filter(!is.infinite(resistance.div))%>%
         group_by(Cno)%>%
         #making quantiles for each year
         dplyr::mutate(
         quantiles_div=cut(resistance.div,
         breaks=quantile(resistance.div, 
         #probs=c(0.25, 0.5, 0.75, 1), 
         probs=seq(0, 1, 0.25),
         na.rm=T),
         labels=c("1", "2", "3", "4")))

#merge with tree attributes

resistance_top4<-merge(resistance_top4, select(trees, -"SPCODE.UPDATE"), by="Tag", all.x=T)

@@

quantiles_2010_top4<-quantile(resistance_top4$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)
quantiles_2015_top4<-quantile(resistance_2015_top4$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

@@

resistance_div<-ggplot(data = resistance_top4, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  facet_grid(.~yr)+
  xlim(-5, 5)+
  xlab("inc/long-term increment")+
  geom_vline(xintercept = c(0,1), lty=2)+
  theme_bw()

@@

resistance_div_sp<-ggplot(data = resistance_top4, 
aes(x=resistance.div, 
col=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_density(alpha=0.5) +
  facet_grid(factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))~yr)+
  scale_color_viridis_d()+ scale_fill_viridis_d()+
  guides(col="none", fill="none")+
  xlim(-5, 5)+
  xlab("inc/long-term increment")+
  geom_vline(xintercept = c(0,1), lty=2)+
  theme_bw()

@@

resistance_map<-ggplot(resistance_top4, aes(x=X, y=Y, size=as.numeric(DBH.04), col=quantiles_div))+
  geom_point(alpha=0.5)+
   scale_size(range = c(.1, 10), name="DBH (cm)")+
  guides(color=guide_legend(title="resistance"))+
  scale_colour_gradient(low="red", high="blue") +
  facet_grid(sp~yr)+
  theme_minimal()+
  coord_fixed()
